Thesis Title

### 2022-09-05 chapters/discussion.tex
Updates to Discussion

The analysis of biological data is driven by the development of a myriad of open-source software tools, each carrying out a specialised step, that then can be chained together to generate new knowledge and insights. However, as with any complex system, susceptibility to variation can cause the entire system to collapse. Such variation can be the use of different operating systems, the availability of computational resources, and ambiguities with tool versioning and documentation \citep{wratten_reproducible_2021}.

@@

In Chapters \ref{ch:paper4} and \ref{ch:paper5}, one of the main key objectives was to a fully reproducible, scalable workflow for one, the analysis and characterization of \ac{DENV} directly from metagenomic samples, and two, evaluate the metagenomic \textit{de novo} assembly process through the benchmark of the most commonly used tools. For that, several steps were implemented.

@@

Favouring open-source tools, with clear documentation describing the methodology implemented, and stating the version of the software used and which parameters were used enables the comparison of results. This requires good management of the software installation, and its dependencies, for this information to be trackable and distributed to the community. Bioinformatics software comes in a variety of programming languages and requires diverse installation methods. This heterogeneity makes the management of software complicated, error-prone, and time-consuming. To reproduce the same software on different machines, full control of the computational environment is required, which usually requires administrative privileges that are often not available.

@@

Remote repositories, such as GitHub\footnote{\url{https://github.com/}}, the single largest host for Git repositories, are the central point of collaboration for millions of developers and projects. A large percentage of all Git repositories are hosted on GitHub, and many open-source projects use it for Git hosting, issue tracking, code review, and code dissemination.

\subsubsection{The need for Open Integration Testing }

>>>>

The analysis of biological data is driven by the development of a myriad of open-source software tools, each carrying out a specialised step, that then can be chained together to generate new knowledge and insights. However, as with any complex system, susceptibility to variation can cause the entire system to collapse. Such variation can be the use of different operating systems, the availability of computational resources, and ambiguities with tool versioning and documentation \citep{wratten_reproducible_2021, mangul_challenges_2019}. No single strategy is sufficient for every scenario, therefore the combination of a multitude of different approaches is necessary to ensure software reproducibility.

@@

In Chapters \ref{ch:paper4} and \ref{ch:paper5}, one of the main key objectives was to a fully reproducible, scalable workflow for one, the analysis and characterization of \ac{DENV} directly from metagenomic samples, and two, evaluate the metagenomic \textit{de novo} assembly process through the benchmark of the most commonly used tools. For that, several steps were implemented. Below there are several strategies that have been implemented throughout this work to ensure reproducibility of the results obtained.

\subsubsection{The need for clear documentation}

The most fundamental strategy for enabling the reproducibility of software, and the results obtain thereafter, is through clear documentation of both the software and the process followed. This includes not only a clear explanation on how the software works, what is expected as input and as output, but also the inclusion of a "quick start" guide to validate tool installation and behaviour. Additionally, the provision of an example dataset is often invaluable to have a "ground truth" to compare the obtained results against, but this will be covered more in depth in the \ref{sssection:integration_testing} section.

The narratives presented in the documentation, both in the manuscripts describing the use of a given software tool, or in the location where the software is hosted, help the community understand how to re-execute a given analysis, and more importantly, how to apply the same process to their own data. This requires that the steps followed are written in enough detail as to be able to be retraced by others. Despite these efforts, documentation alone is not enough to ensure reproducibility as computer configurations can differ greatly, from different operative systems to the libraries available in the system itself.

As important as the overall narrative describing a tool, clear documentation within the tool itself, added during its development, represents an important step for readability, and therefore accountability, of the code produced. This also represents an important requirement for the maintenance of the tool long-term, even after the original developer or team of developers is no longer actively maintaining the software.

@@

Favouring open-source tools, with clear documentation describing the methodology implemented, and stating the version of the software used and which parameters were used enables the comparison of results. This requires good management of the software installation, and its dependencies, for this information to be trackable and distributed to the community. Bioinformatics software comes in a variety of programming languages and requires diverse installation methods. This heterogeneity makes the management of software complicated, error-prone, and time-consuming. To reproduce the same software on different machines, full control of the computational environment is required, which usually requires administrative privileges that are often not available.  It has been described that as much as 28\% of all resources are currently not accessible through the original \ac{URL} published in the paper they first appeared in \citep{mangul_challenges_2019}. Although several journals have clear requirements for the sharing of code and data, there are currently no effective requirements to promote installability and long-term archival stability of software tools \citep{mangul_challenges_2019}, strongly hindering the effectiveness of the tool distribution and usability.

@@

Remote repositories, such as GitHub\footnote{\url{https://github.com/}}, the single largest host for Git repositories, are the central point of collaboration for millions of developers and projects. A large percentage of all Git repositories are hosted on GitHub, and many open-source projects use it for Git hosting, issue tracking, code review, and code dissemination. This benefits the computational biology community immensely, effectively guaranteeing permanent access to published scientific \ac{URL} \citep{mangul_challenges_2019}, either directly or through an archived version.

\subsubsection{The need for Open Integration Testing} \label{sssection:integration_testing}
